---
title: "Density_Project_Module"
author: "Abby Robinson"
date: "1/30/2023"
output: html_document
---

# How do density dependent dynamics influence the efficacy of mimetic signaling? 

## This module will walk you through conducting a robust statistical analysis of data collected at La Selva Biological Station during the summer of 2022. 

## First step: Connecting to GitHub

You'll need to create a project in RStudio on your computer and link to to my GitHub repository for this project. You can find my GitHub repo [here](https://github.com/butterfliesrcool/Density_Dependent_Dynamics)

### To link a R Project to a GitHub Repo, 

1. In RStudio, go to "file", "new project" 

2. Select "version control" and "git"

3. Input the URL for my repo (above), name your project, and chose where on your computer you want to store it 

4. You can now create your own R Markdown file to begin writing your code!

If you have any issues connecting to GitHub or want a more detailed refresher, check out Dr. Chris Schmitt's GitHub module [here](https://fuzzyatelin.github.io/bioanth-stats/module-02/module-02.html)

## Initial Data Visualization 

### What are the goals of this project? 

we are testing the hypothesis that variation in the absolute density of Batesian models and their imperfect mimics impacts the rate of predation on artificial butterfly facsimiles using a paired experimental design (low vs. high density)

### Loading the data 

At this point, you have used the final sorted attack photos to write two data sets: one with count data and another with binomial data. You should have also calculated, by hand, the attack totals on each day. Please compare those attack rates to the ones calculated below. If they do not match, you may need to check your excel file for spaces using the find and replace functions 

Install the "Curl" Package and then load the package library using the code below 

```{r}
library(curl)
```

Use the "curl" command to load the data into R using the data file link from the repo 

```{r}
c <- curl("https://raw.githubusercontent.com/butterfliesrcool/Temporal_Dynamics/main/Quabbin_Mimic_Data.csv")
mimic <- read.csv(c, header = TRUE, sep = ",")
head(mimic)
```

Making sure R is "reading" the variables field.day, experiment.day, field.site, species, and treatment at factors. Attacks variable should be an integer value 

```{r}
# make experimental.day variable factor 
class(mimic$experiment.day) 
mimic$experiment.day <- as.factor(mimic$experiment.day)
class(mimic$experiment.day)

# make field.day variable factor
class(mimic$field.day) 
mimic$field.day <- as.factor(mimic$field.day)
class(mimic$field.day)

# make site variable factor 
class(mimic$field.site)
mimic$field.site <- as.factor(mimic$field.site)
class(mimic$field.site)

# make sure attack variable is integer 
class(mimic$attacks) # only thing that should be an integer 

# make species variable factor 
class(mimic$species)
mimic$species <- as.factor(mimic$species)
class(mimic$species)

# make treatment variable factor 
class(mimic$treatment)
mimic$treatment <- as.factor(mimic$treatment)
class(mimic$treatment)
```

```{r}
# histogram of attacks 
hist(mimic$attacks)  #concerns about zero inflation based on histogram? 

# check variance and mean to assess for over /  under - dispersion 
var(mimic$attacks)
mean(mimic$attacks) # variance is greater than mean, which indicates that over-dispersion might be an issue 
```

```{r}
# summarizing attacks 
aggregate(x= mimic$attacks, by= list(mimic$experiment.day, mimic$species), FUN=sum)
```

Do these numbers match yours? If so, we can move on! 

## Model Building 

Install and load  the lme4, glmmTMB, and performance packages 

```{r}
library(lme4)
library("glmmTMB")
library(performance)
```

Here we start the process of building a generalized linear mixed model (GLMM). The model below was used to analyze the data from the quabbin reservoir. We will modify it to fit the data and research questions from Costa Rica. First, our response variable is "attacks", when this data is recorded as counts (total number of attacks on each species), we can use the Poisson distribution (family=poisson). However, the assumptions of the Poisson distribution may not be met if our data is zero inflated. If this happened, we can try to adjust the model or use the negative binomial distribution instead. 

the * in the model below indicates an interaction between variables. For example, (attacks ~ treatment + species) would tell you if attacks vary between different treatments and if attacks vary between species, independently. Alternatively, (attacks ~ treatment*species) would tell you if the total number of attacks varies across species within different treatments. We will need to figure out the right coding syntax to address our hypothesis from the Costa Rica project. 

The variables discussed above are fixed variables because they represent specific treatments that we employed. The location of field sites and the day that we collected data are random effect variables because they are things that could influence our results, but were not specifically controled in the experiment. 

```{r}
mod1 <- glmmTMB(attacks ~ experiment.day*species + treatment*species + (1|field.site) + (1|field.day), data=mimic,ziformula=~1,family=poisson) 
summary(mod1)
check_singularity(mod1) 

# singularity issues, likely due to the field.day random effect variable 
```

We will need to check this model for singularity issues, convergence issues, error messages, residuals, and assumptions fit 

Check residuals with "DHARMa" package 

```{r}
library(DHARMa)
```

We used the "DHARMa" package to check residuals and model assumptions 
```{r}
vignette("DHARMa", package="DHARMa") ### Useful information for checking residuals/ model assumptions 
```

```{r}
mod1simulation <- simulateResiduals(fittedModel = mod1, plot = T)
testUniformity(mod1simulation) 
testOutliers(mod1simulation)
testDispersion(mod1simulation) 
testZeroInflation(mod1simulation) 
```

## Post Hoc Pairwise Comparison 

```{r}
library(emmeans)
```

For the Costa Rica project, we want to compare attacks on iphiclus and serpa directly to attacks on junonia to account for variation in the number of facsimiles in the low density and high density experiments. To do this, we will perform a pairwise post hoc test comparing species within the different treatments 

```{r}
post.hoc1 <- emmeans(mod1, pairwise ~ species|treatment) 
#there is a significant difference in the attack rates on junonia and limenitis only in the zero week treatment (simultaneous presentation)
post.hoc1
```


## Summary 

This is the analysis you will have to modify and apply to the Costa Rica Project 

